Data Exploration: IUCN and AquaMaps

### 2015-09-18 data_explore.Rmd
added ggsave for plots throughout

print(g)
```

@@

dot = ggplot(df,aes(iucn_area,am_area,color=category))+ geom_point() + geom_abline()+
      labs(x='Area (km2) from IUCN dataset',y='Area (km2) from AquaMaps dataset',title='Comparing total area of species range maps')
print(dot)
```

@@

all_status <- ggplot(df,aes(iucn_log,am_log,color=category)) + 
  geom_point() + geom_abline() +
  labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
        title= 'Comparing area of range maps\n (all area values are km2 log transformed)',fill='IUCN Category')

print(all_status)
```

@@

groups <- ggplot(df, aes(iucn_log, am_log,color=category)) + geom_point() +facet_wrap(~spp_group) + 
        geom_abline() + labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
                             title='Comparing differences in area across species groups', fill='IUCN Category')
print(groups)
```
***

@@

all<-ggplot(overlap,aes(percent))+
    geom_histogram(aes(y=..count../1926),binwidth=5,alpha=.8)+
    labs(x='Percent of total range that is shared between both datasets', y='Proportion of Total Species')

print(all)
```

@@

sp <- ggplot(overlap,aes(percent))+geom_histogram(binwidth=5,alpha=.8) +
  facet_wrap(~spp_group) +
  labs(x='Percent of total range that is shared between both datasets', y='Number of Total Species')

print(sp)

@@

sm_plot <- ggplot(d, aes(sm_perc, fill = sm_range)) + 
       geom_histogram(aes(y=..count../sum(..count..)),binwidth = 4, alpha = .8) +
        scale_fill_discrete(name = "Dataset with smaller range", labels = c("AquaMaps", "IUCN")) + 
        labs(x = 'Percent of smaller range that falls within the larger range', y = 'Proportion of all species (n=1962)')

@@

@@

@@

@@

@@

  qtile_plot <- ggplot(area_clip, aes(total_area, sm_perc, color = sm_range)) + 
  geom_point() + 
  labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
       title = sprintf('Size of species range (quartile %s) vs. percent overlap of \nsmaller range with larger', i))

print(qtile_plot)
}
# First quartile regression:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept)  1423917     101053  14.091   <2e-16 ***
#     sm_perc         1188       1555   0.764    0.445    
#     Multiple R-squared:  0.001198,	Adjusted R-squared:  -0.0008529 
#
# second quartile:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept)  7011028     275943  25.408  < 2e-16 ***
#     sm_perc       -17725       4228  -4.192 3.28e-05 ***
#     Multiple R-squared:  0.03483,	Adjusted R-squared:  0.03285 
#
# third quartile:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept) 12813279     507856  25.230  < 2e-16 ***
#     sm_perc        64710       8159   7.931  1.5e-14 ***
#     Multiple R-squared:  0.1146,	Adjusted R-squared:  0.1128 
#
# fourth quartile:
#                   Estimate Std. Error t value Pr(>|t|)    
#     (Intercept) -101736351   17259544  -5.894 7.04e-09 ***
#     sm_perc        1911286     215216   8.881  < 2e-16 ***
#     Multiple R-squared:  0.1399,	Adjusted R-squared:  0.1381

@@

hist <- ggplot(df,aes(prop,fill=iucn_code))+geom_histogram(aes(y=..count../sum(..count..)),binwidth = 0.05, alpha = .7)+labs(x='Proportion of range area included when selecting a threshold of 0.4',y='Proportion of all aquamaps species (17,348 total)')
print(hist)

>>>>

ggsave('figures/sp_groups_barchart.png',plot=g,width=8, height=4, dpi=100)
print(g)

@@

dot = ggplot(df,aes(iucn_area,am_area,color=category))+ geom_point() + geom_abline()+
      labs(x='Area (km2) from IUCN dataset',y='Area (km2) from AquaMaps dataset',title='Comparing total area of species range maps')
ggsave('figures/CompareTotalArea_AM-IUCN_scatter.png',plot=dot,width=8, height=4, dpi=100)
print(dot)
```

@@

all_status <- ggplot(df,aes(iucn_log,am_log,color=category)) + 
  geom_point() + geom_abline() +
  labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
        title= 'Comparing area of range maps\n (all area values are km2 log transformed)',fill='IUCN Category')
ggsave('figures/CompareTotalAreaLOG_AM-IUCN_scatter.png',plot=all_status,width=8,height=4,dpi=100)
print(all_status)
```

@@

groups <- ggplot(df, aes(iucn_log, am_log,color=category)) + geom_point() +facet_wrap(~spp_group) + 
        geom_abline() + labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
                             title='Comparing differences in area across species groups', fill='IUCN Category')
ggsave('figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png',plot=groups,width=8,height=6,dpi=100)
print(groups)
```
***

@@

all<-ggplot(overlap,aes(percent))+
    geom_histogram(aes(y=..count../1926),binwidth=5,alpha=.8)+
    labs(x='% total range that is shared b/w both datasets', y='Proportion of Total Species')
ggsave('figures/Overlap-PercTotalRange.png')
print(all)
```

@@

sp <- ggplot(overlap,aes(percent))+geom_histogram(binwidth=5,alpha=.8) +
  facet_wrap(~spp_group) +
  labs(x='Percent of total range that is shared between both datasets', y='Number of Total Species')+ 
       theme(strip.text.x = element_text(size = 8))

ggsave('figures/Overlap-PercTotalRange_spGroups.png')
print(sp)

@@

sm_plot <- ggplot(d, aes(sm_perc, fill = sm_range)) + 
       geom_histogram(aes(y=..count../sum(..count..)),binwidth = 4, alpha = .8) +
        scale_fill_discrete(name = "Dataset with smaller range", labels = c("AquaMaps", "IUCN")) + 
        labs(x = 'Percent of smaller range that falls within the larger range', y = 'Proportion of all species (n=1962)')
ggsave('figures/PercSmallRangeInLarger.png',width=6,height=4)

@@

ggsave('figures/AreavsPercOverlap.png',width=6,height=4)

@@

ggsave('figures/CorrelationAreavsPercOverlap.png')

@@

ggsave('figures/AreavsPercOverlap_2.png',width=6,height=4)

@@

ggsave('AreavsPercOverlap_log.png',width=6,height=4)

@@

  qtile_plot <- ggplot(area_clip, aes(total_area, sm_perc, color = sm_range)) + 
  geom_point() + 
  labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
       title = sprintf('Size of species range (quartile %s) vs. percent overlap of \nsmaller range with larger', i))
ggsave(paste0('figures/AreavsOverlap_quantile_',i,'.png'),width=6,height=4)
print(qtile_plot)
}
# First quartile regression:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept)  1423917     101053  14.091   <2e-16 ***
#     sm_perc         1188       1555   0.764    0.445    
#     Multiple R-squared:  0.001198,	Adjusted R-squared:  -0.0008529 
#
# second quartile:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept)  7011028     275943  25.408  < 2e-16 ***
#     sm_perc       -17725       4228  -4.192 3.28e-05 ***
#     Multiple R-squared:  0.03483,	Adjusted R-squared:  0.03285 
#
# third quartile:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept) 12813279     507856  25.230  < 2e-16 ***
#     sm_perc        64710       8159   7.931  1.5e-14 ***
#     Multiple R-squared:  0.1146,	Adjusted R-squared:  0.1128 
#
# fourth quartile:
#                   Estimate Std. Error t value Pr(>|t|)    
#     (Intercept) -101736351   17259544  -5.894 7.04e-09 ***
#     sm_perc        1911286     215216   8.881  < 2e-16 ***
#     Multiple R-squared:  0.1399,	Adjusted R-squared:  0.1381

@@

hist <- ggplot(df,aes(prop,fill=iucn_code))+geom_histogram(aes(y=..count../sum(..count..)),binwidth = 0.05, alpha = .7)+labs(x='Proportion of range area included when selecting a threshold of 0.4',y='Proportion of all aquamaps species (17,348 total)')
ggsave('figures/AMThreshold.png',width=6,height=4)
print(hist)
